Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-03-01 article_springer/article_springer.Rmd
Cleaned Up
First changes (addressing Reviewer 2)

bibliography: CTRMstats.bib
csl: 
output: 
       rticles::elsevier_article:
                fig_caption: yes
                number_sections: true

@@

Often a magnitude, or mark can be assigned to each event in the renewal process, 
such as for earthquakes, solar flares or neuron voltages. 
The Peaks-Over-Threshold model [POT, see e.g. @ColesBook] applies a
threshold to the magnitudes, and fits a Generalized Pareto distribution 
to the threshold exceedances. 
A commonly made assumption in POT models is that times between 
events are either fixed 
or light-tailed, and this entails that the 
threshold crossing times form a Poisson process [@Hsing88]. 
Then as one increases the threshold and thus decreases the threshold 
crossing probability $p$, the Poisson process is rarefied, i.e. its 
intensity decreases _linearly_ with $p$ [see e.g. @beirlantBook].

As will be shown below, in the heavy-tailed waiting time scenario threshold 
crossing times form a _fractional Poisson process_ 
[@Laskin2003; @Meerschaert2010b], which is a 
renewal process with Mittag-Leffler distributed waiting times. 
The family of Mittag-Leffler distributions nests the exponential 
distribution [@Haubold11], and hence the fractional Poisson process
generalizes the standard Poisson process. 
Again as the threshold size increases and the threshold crossing
probability $p$ decreases, the fractional Poisson process is rarefied: 
The scale parameter of the Mittag-Leffler inter-arrival times of 
threshold crossing times increases, but _superlinearly_; 
see the Theorem below.

@@

: Let $(W,J), (W_1, J_1), (W_2, J_2), \ldots$ be i.i.d. pairs of random 
variables, where the $W_k > 0$ are interpreted as the *waiting times* 
and $J_k \in [x_L, x_R]$ as the *event magnitudes* 
($x_L \in [-\infty, +\infty), x_R \in (-\infty, +\infty]$). 
If $W$ and $J$ are independent, the Marked Renewal Process is said to be *uncoupled*.

We assume that the $k$-th magnitude $J_k$ occurs at time
$T_k = W_1 + \ldots + W_k$. 
Based on an MRP, we define the Continuous Time Random Exceedance model
(CTRE) as follows:

@@

: Given a threshold $\ell \in (x_L, x_R)$, 
consider the stopping time 
$$\tau(\ell) := \min\{k: J_k > \ell\},\quad \ell \in (x_L, x_R).$$
Define the pair of random variables $(X(\ell), T(\ell))$ via 
$$X(\ell) = J_{\tau(\ell)} - \ell, \quad 
T(\ell) = \sum_{k=1}^{\tau(\ell)} W_k.$$
By restarting the MRP at $\tau(\ell)$, inductively define the
two i.i.d. sequences $T(\ell,n)$ and $X(\ell, n)$, $n \in \mathbb N$, 
called the "inter-arrival times" and the "exceedances", respectively.
The pair sequence $(T(\ell, n), W(\ell, n))_{n \in \mathbb N}$ is called 
a Continuous Time Random Exceedance model (CTRE). 
If the underlying MRP is uncoupled, then the CTRE is also called 
uncoupled.

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J - \ell | J > \ell$, independent of any waiting time $W_k$.  
Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top panel, where $W$ has a stable distribution with tail parameter $\beta =$ `r params$tail` (and skewness $1$ and location $0$), and where $J$ is from a standard Gumbel distribution. In the bottom panel, we plot a time series of solar flare intensities derived from a NASA dataset [@HXRBS] which we will later examine more closely (see Section 7). Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear continuously. The threshold exceedances, however, appear to have visually similar statistical behaviour in both models. Observations below a threshold are commonly discarded in Extreme Value Theory (POT approach); likewise, the CTRE model interprets these observations as noise and discards them.

@@

: Let the waiting times $W_k$ be in the domain of attraction of a positively skewed sum-stable law with stability parameter $0 < \beta < 1$; more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. Write $p := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p)} \to W_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $W_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sW_\beta)] = \frac{1}{1+s^\beta}.
$$

@@

Without loss of generality, we choose $\ell^*$ such that $F(\ell^*) = 1/e$, 
and let $\ell = a(\tilde b(c)) \ell^* + d(\tilde b(c))$. 
We may then calculate 
$$
\mathbf P\left[ \frac{T(\ell)}{b(1/p)} > t \right]
= \mathbf P[T(\ell) > b(1/p) t]
= \mathbf P[V(ct) \le \ell]
$$
where we have substituted $c = b(1/p)$. Moreover 
$$
\mathbf P[V(ct) \le \ell]
= \mathbf P\left[ \frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} 
\le \frac{\ell - d(\tilde b(c))}{a(\tilde b(c))} \right]
\longrightarrow \mathbf P\left[ Y(t) \le \ell^* \right]
$$
Defining the hitting time of level $\ell^*$ by $Y(t)$ as 
$\xi_{\ell^*} := \inf\{t: Y(t) > \ell^*\}$,
we then have 
$$
P\left[ Y(t) \le \ell^* \right] = \mathbf P[\xi_{\ell^*} > t] 
= \mathbf P[(-\log F(\ell^*))^{-1/\beta} X^{1/\beta} D > t]
$$
by [Proposition 4.2, @MeerschaertStoev08], where $X$ is an exponential 
random variable with mean $1$. 
Using [Theorem 19.1, @Haubold11], we see that 
$X^{1/\beta} D \sim {\rm ML}(\beta, 1)$, concluding the proof. 
\qed

For a scale parameter $\sigma > 0$, we write ${\rm ML}(\beta, \sigma)$ for the distribution of $\sigma W_\beta$. The Mittag-Leffler distribution with parameter $\beta \in (0,1]$ is 
a heavy-tailed positive distribution for $\beta < 1$, with infinite mean. 
However, as $\beta \uparrow 1$, ${\rm ML}(\beta, \sigma)$ converges 
weakly to the exponential distribution ${\rm Exp}(\sigma)$ with mean $\sigma$.
This means that although its moments are all infinite, the Mittag-Leffler
distribution may (if $\beta$ is close to 1) be indistinguishable from the 
exponential distribution, for the purposes of applied statistics.

@@

The classical POT approach is based on the fact that exceedances above a high threshhold are asymptotically GPD distributed, hence a GPD distribution is fitted to the exceedances above a high enough threshhold. Due to the Theorem in the last Section we know that for high thresholds the times between exceedances are asymptotically Mittag-Leffler distributed in case of heavy-tailed waiting times between the observations. Hence, before we talk about inference for the exceedances in the next section, we discuss inference for Mittag-Leffler distributions.

@@

The Theorem in Section \ref{sec:scaling} implies that for a high 
threshold $\ell$ we may approximate the distribution of $T(\ell)$ with an 
${\rm ML}(\beta, b(1/p))$ distribution, 
where the function $b(c)$ varies regularly at $\infty$ with parameter $1/\beta$.
Building on the POT (Peaks-Over-Threshold) method, we propose the following estimation procedure for the distribution of inter-exceedance time $T(\ell)$:

@@

To clarify Step 4: Recall that by the theorem, $\sigma_k / b(1/p)$ is expected to stabilize 
around a constant as $k \downarrow K_0$. Since $b$ is regularly varying with parameter
$1/\beta$, we have $b(1/p) = p^{-1/\beta} / L(1/p)$ for some slowly varying function $L$. 
Approximating $p \approx \hat p := k / n$, we have 
$$\text{const} \approx \sigma_k / b(1/p) = \sigma_k \times p^{1/\beta} L(1/p) \approx \sigma_k \times k^{1/\beta} n^{-1/\beta} L(n/k)$$
Assuming that the variation of $L(n/k)$ is minor, we can hence see that 
$\sigma_k k^{1/\beta}$ stabilizes.

@@

**Case (i):** For (i) we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation [see e.g. the vignette on parametrisation in @MittagLeffleR]
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist. By the Theorem, the distribution of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$.

@@

Figure \ref{fig:flares} shows the stability plots for the solar flare data, on the left for the tail parameter and on the right for the scale parameter. The dark grey ranges correspond to 95% confidence intervals, which are derived from the asymptotic normality of the log-moments estimators [@Cahoy2013] and the $\delta$-method [@MittagLeffleR]; dashed lines show the deduced true values of $\beta$ resp.\ $\sigma_0$. The stability plot for the tail stabilizes nicely around 0.85 (dashed line), while 
the scale parameter stabilizes less obviously near $3 \times 10^7$ (dashed line).
The growth of the scale parameter for lower threshold appears to be closer to linear in $p$, 
rather than proportional to $p^{1/0.85}$ as suggested by the Mittag-Leffler
fits. The reason for this is likely that the overall goodness of fit as 
compared to an exponential distribution is improved due to the 
peaked shape of the Mittag-Leffler distribution near $0$,
rather than its tail behaviour at $\infty$.
The reported fit should hence come with the caveat that a Mittag-Leffler 
distribution models exceedance times well only up to certain time-scales. 
More research is needed into the modelling of scale transitions, where
inter-exceedance times appear to have different power laws across different 
time scales.

@@

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the estimated threshold exceedance time distribution is 
$$
T(\ell) \sim {\rm ML}(\beta, k^{-1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not memoryless, and the probability density of the time $t$ until the next threshold crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{-1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{-1/\beta} \sigma_0)$ is the probability density of ${\rm ML}(\beta, k^{-1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{-1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{-1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a decreasing function for the Mittag-Leffler distribution. 
```{r hazard, fig.height=4, fig.cap="\\label{fig:hazard} Left: Conditional distribution of time until next threshold crossing, depending on elapsed time $t_0$ since last crossing ($\\beta = 0.8$, $\\sigma_0 = 1$). Right: Hazard rate depending on tail parameter $\\beta$."}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(1,2))
df1<-data.frame(x=xx,y=rep(1, length(xx)))
col<-c("red","blue","darkgreen","black")

@@

We have extended the POT (Peaks-over-Threshold) model, a mainstay of extreme value theory, to "bursty" time series, which have been studied 
intensively in statistical physics. Burstiness is characterized by power-law waiting times between events, and we have shown that the Mittag-Leffler distribution arises naturally as a scaling limit for the inter-exceedance times of high thresholds. Moreover, we have derived the following non-linear scaling behaviour: $\sigma \sim p^{-1/\beta}$, where $\sigma$ is the scale parameter of the distribution of threshold 
exceedance times, $p$ is the fraction of magnitudes above the threshold, and $\beta$ the exponent of the power law. 
This "anomalous" scaling behaviour in the bursty setting entails two phenomena:

>>>>

bibliography: CTRMstats.bib
csl: elsevier-harvard.csl
output: 
  rticles::elsevier_article:
    number_sections: true
    fig_caption: yes

@@

Often a magnitude, or mark can be assigned to each event in the renewal process, 
such as for earthquakes, solar flares or neuron voltages. 
The Peaks-Over-Threshold model [POT, see e.g. @ColesBook] applies a
threshold to the magnitudes, and fits a Generalized Pareto distribution 
to the threshold exceedances. 
A commonly made assumption in POT models is that times between 
events are either fixed 
or light-tailed, and this entails that the 
threshold crossing times form a Poisson process [@Hsing88]. 
Then as one increases the threshold and thus decreases the threshold 
crossing probability $p_{\ell}$, the Poisson process is rarefied, i.e. its 
intensity decreases _linearly_ with $p_{\ell}$ [see e.g. @beirlantBook].

As will be shown below, in the heavy-tailed waiting time scenario threshold 
crossing times form a _fractional Poisson process_ 
[@Laskin2003; @Meerschaert2010b], which is a 
renewal process with Mittag-Leffler distributed waiting times. 
The family of Mittag-Leffler distributions nests the exponential 
distribution [@Haubold11], and hence the fractional Poisson process
generalizes the standard Poisson process. 
Again as the threshold size increases and the threshold crossing
probability $p_{\ell}$ decreases, the fractional Poisson process is rarefied: 
The scale parameter of the Mittag-Leffler inter-arrival times of 
threshold crossing times increases, but _superlinearly_; 
see the Theorem below.

@@

: Let $(W,J), (W_1, J_1), (W_2, J_2), \ldots$ be i.i.d. pairs of random variables, where the $W_k > 0$ are interpreted as the *waiting times* 
and $J_k \in \mathbb{R}$ as the *event magnitudes*. If $W$ and $J$ are independent, the Marked Renewal Process is said to be *uncoupled*.

In the following we denote with $x_L \in [-\infty, +\infty)$ and $x_R \in (-\infty, +\infty]$) the left and right endpoint of the distribution of $J$. We assume that the $k$-th magnitude $J_k$ occurs at time
$T_k = W_1 + \ldots + W_k$. 
Based on an MRP, we define the Continuous Time Random Exceedance model
(CTRE) as follows:

@@

: Given a threshold $\ell \in (x_L, x_R)$, 
consider the stopping time 
$$\tau(\ell) := \min\{k: J_k > \ell\},\quad \ell \in (x_L, x_R).$$
Define the pair of random variables $(X(\ell), T(\ell))$ via 
$$X(\ell) = J_{\tau(\ell)} - \ell, \quad 
T(\ell) = \sum_{k=1}^{\tau(\ell)} W_k.$$
By restarting the MRP at $\tau(\ell)$, inductively define the
two i.i.d. sequences $T(\ell,n)$ and $X(\ell, n)$, $n \in \mathbb N$, 
called the "inter-arrival times" and the "exceedances", respectively.
The pair sequence $(X(\ell, n), T(\ell, n))_{n \in \mathbb N}$ is called 
a Continuous Time Random Exceedance model (CTRE). 
If the underlying MRP is uncoupled, then the CTRE is also called 
uncoupled.

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J - \ell | J > \ell$, independent of any waiting time $W_k$.  
We assume for the rest of the article, that the magnitudes $(J_i)_{i \in \mathbb{N}}$ belong to the the max-domain of attraction of some non-degenerate distribution. This means there exist $a_n>0$ and $d_n \in \mathbb{R}$ such that 
$$ a_n^{-1}(J_1 \vee \ldots \vee J_n -d_n) \Rightarrow A \text{ as } n \rightarrow \infty.$$
Hence, the distribution of $A$ is a generalized extreme value distribution (GEV) which distribution function is given by 
$$ F(x;\xi) = \begin{cases}\exp(-(1+\xi x)^{-1/\xi}) & \xi\neq0 \\ \exp(-\exp(-x)) & \xi = 0\end{cases}.$$ 
The GEV is subdivided into the Gumbel ($\xi=0$), the Weibull ($\xi<0$) and the Fr'echet ($\xi>0$) family of distributions.

Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top panel, where $W$ has a stable distribution with tail parameter $\beta =$ `r params$tail` (and skewness $1$ and location $0$), and where $J$ is from a standard Gumbel distribution. In the bottom panel, we plot a time series of solar flare intensities derived from a NASA dataset [@HXRBS] which we will later examine more closely (see Section 7). Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear continuously. The threshold exceedances, however, appear to have visually similar statistical behaviour in both models. Observations below a threshold are commonly discarded in Extreme Value Theory (POT approach); likewise, the CTRE model interprets these observations as noise and discards them.

@@

: Let the waiting times $W_k$ be in the domain of attraction of a positively skewed sum-stable law with stability parameter $0 < \beta < 1$; more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. Write $p_{\ell} := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p_{\ell})} \to Z_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $Z_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sZ_\beta)] = \frac{1}{1+s^\beta}.
$$

@@

Without loss of generality, we choose $\ell^*$ such that $F(\ell^*) = 1/e$, 
and let $\ell = a(\tilde b(c)) \ell^* + d(\tilde b(c))$. 
We may then calculate 
$$
\mathbf P\left[ \frac{T(\ell)}{b(1/p_{\ell})} > t \right]
= \mathbf P[T(\ell) > b(1/p_{\ell}) t]
= \mathbf P[V(ct) \le \ell]
$$
where we have substituted $c = b(1/p_{\ell})$. Moreover 
$$
\mathbf P[V(ct) \le \ell]
= \mathbf P\left[ \frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} 
\le \frac{\ell - d(\tilde b(c))}{a(\tilde b(c))} \right]
\longrightarrow \mathbf P\left[ Y(t) \le \ell^* \right]
$$
Defining the hitting time of level $\ell^*$ by $Y(t)$ as 
$\xi_{\ell^*} := \inf\{t: Y(t) > \ell^*\}$,
we then have 
$$
P\left[ Y(t) \le \ell^* \right] = \mathbf P[\xi_{\ell^*} > t] 
= \mathbf P[(-\log F(\ell^*))^{-1/\beta} X^{1/\beta} D > t]
$$
by [Proposition 4.2, @MeerschaertStoev08], where $X$ is an exponential 
random variable with mean $1$. 
Using [Theorem 19.1, @Haubold11], we see that 
$X^{1/\beta} D \sim {\rm ML}(\beta, 1)$, concluding the proof. 
\qed

For a scale parameter $\sigma > 0$, we write ${\rm ML}(\beta, \sigma)$ for the distribution of $\sigma Z_\beta$. The Mittag-Leffler distribution with parameter $\beta \in (0,1]$ is 
a heavy-tailed positive distribution for $\beta < 1$, with infinite mean. 
However, as $\beta \uparrow 1$, ${\rm ML}(\beta, \sigma)$ converges 
weakly to the exponential distribution ${\rm Exp}(\sigma)$ with mean $\sigma$.
This means that although its moments are all infinite, the Mittag-Leffler
distribution may (if $\beta$ is close to 1) be indistinguishable from the 
exponential distribution, for the purposes of applied statistics.

@@

The classical POT approach is based on the fact that exceedances above a high threshold are asymptotically GPD distributed, hence a GPD distribution is fitted to the exceedances above a high enough threshold. Due to the Theorem in the last Section we know that for high thresholds the times between exceedances are asymptotically Mittag-Leffler distributed in case of heavy-tailed waiting times between the observations. Hence, before we talk about inference for the exceedances in the next section, we discuss inference for Mittag-Leffler distributions.

@@

The Theorem in Section \ref{sec:scaling} implies that for a high 
threshold $\ell$ we may approximate the distribution of $T(\ell)$ with an 
${\rm ML}(\beta, b(1/p_{\ell}))$ distribution, 
where the function $b(c)$ varies regularly at $\infty$ with parameter $1/\beta$.
Building on the POT (Peaks-Over-Threshold) method, we propose the following estimation procedure for the distribution of inter-exceedance time $T(\ell)$:

@@

To clarify Step 4: Recall that by the theorem, $\sigma_k / b(1/p_{\ell})$ is expected to stabilize 
around a constant as $k \downarrow K_0$. Since $b$ is regularly varying with parameter
$1/\beta$, we have $b(1/p_{\ell}) = p_{\ell}^{-1/\beta} / L(1/p_{\ell})$ for some slowly varying function $L$. 
Approximating $p_{\ell} \approx \hat p_{\ell} := k / n$, we have 
$$\text{const} \approx \sigma_k / b(1/p_{\ell}) = \sigma_k \times p_{\ell}^{1/\beta} L(1/p_{\ell}) \approx \sigma_k \times k^{1/\beta} n^{-1/\beta} L(n/k)$$
Assuming that the variation of $L(n/k)$ is minor, we can hence see that 
$\sigma_k k^{1/\beta}$ stabilizes.

@@

**Case (i):** For (i) we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation [see e.g. the vignette on parametrisation in @MittagLeffleR]
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist. By the Theorem, the distribution of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p_{\ell}^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$.

@@

Figure \ref{fig:flares} shows the stability plots for the solar flare data, on the left for the tail parameter and on the right for the scale parameter. The dark grey ranges correspond to 95% confidence intervals, which are derived from the asymptotic normality of the log-moments estimators [@Cahoy2013] and the $\delta$-method [@MittagLeffleR]; dashed lines show the deduced true values of $\beta$ resp.\ $\sigma_0$. The stability plot for the tail stabilizes nicely around 0.85 (dashed line), while 
the scale parameter stabilizes less obviously near $3 \times 10^7$ (dashed line).
The growth of the scale parameter for lower threshold appears to be closer to linear in $p_{\ell}$, 
rather than proportional to $p_{\ell}^{1/0.85}$ as suggested by the Mittag-Leffler
fits. The reason for this is likely that the overall goodness of fit as 
compared to an exponential distribution is improved due to the 
peaked shape of the Mittag-Leffler distribution near $0$,
rather than its tail behaviour at $\infty$.
The reported fit should hence come with the caveat that a Mittag-Leffler 
distribution models exceedance times well only up to certain time-scales. 
More research is needed into the modelling of scale transitions, where
inter-exceedance times appear to have different power laws across different 
time scales.

@@

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the estimated threshold exceedance time distribution is 
$$
T(\ell) \sim {\rm ML}(\beta, k^{-1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not memoryless, and the probability density of the time $t$ until the next threshold crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density is approximately equal to 
$$
p(t|\beta, \sigma_0, \ell, t_0) \approx \frac{f(t + t_0 | \beta, k^{-1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{-1/\beta} \sigma_0)$ is the probability density of ${\rm ML}(\beta, k^{-1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{-1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{-1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a decreasing function for the Mittag-Leffler distribution. 
```{r hazard, fig.height=4, fig.cap="\\label{fig:hazard} Left: Conditional distribution of time until next threshold crossing, depending on elapsed time $t_0$ since last crossing ($\\beta = 0.8$, $\\sigma_0 = 1$). Right: Hazard rate depending on tail parameter $\\beta$."}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(1,2))
df1<-data.frame(x=xx,y=rep(1, length(xx)))
col<-c("red","blue","darkgreen","black")

@@

We have extended the POT (Peaks-over-Threshold) model, a mainstay of extreme value theory, to "bursty" time series, which have been studied 
intensively in statistical physics. Burstiness is characterized by power-law waiting times between events, and we have shown that the Mittag-Leffler distribution arises naturally as a scaling limit for the inter-exceedance times of high thresholds. Moreover, we have derived the following non-linear scaling behaviour: $\sigma \sim p_{\ell}^{-1/\beta}$, where $\sigma$ is the scale parameter of the distribution of threshold 
exceedance times, $p_{\ell}$ is the fraction of magnitudes above the threshold, and $\beta$ the exponent of the power law. 
This "anomalous" scaling behaviour in the bursty setting entails two phenomena:
